alternativeRSA

### 2016-01-18 models/RSA_alternativeCostFormulation.Rmd
correct speaker.inf3 working. Also fixed normalization in speaker.lhd

New Prior incorporation
```{r}
speaker.inf3 = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd3 = function(currRating, currDegree, m, alpha, costs, allDegrees) {
  currCost <- subset(costs, degree == currDegree)$cost
  numerator = speaker.inf3(m[[currDegree]][currRating], alpha, currCost)
#   normalize = sum(unlist(sapply(seq(nrow(m)),
#                                 function(i) {
#                                   speaker.inf3(m[i, currDegree],
#                                                alpha,
#                                                subset(costs, degree == currDegree)$cost)})))
  # normalize over other words
  normalize <- sum(sapply(seq(allDegrees), FUN = 
                            function(i) {
                              speaker.inf(m[[allDegrees[i]]][currRating], alpha,
                                          subset(costs, degree == allDegrees[i])$cost)}
                          ))
  return(numerator / normalize)
  # return(numerator)
}
# old !!!!
# speaker.lhd = function(rating, degree, m, alpha) {
#   numerator = speaker.inf(m[rating, degree], alpha)
#   normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
#   return(numerator / normalize)
# }

@@

New run fn()
```{r}
computePosteriors <- c(norm.post, norm.post3)
runFull2 <- function(data, calcFn = 1, alpha = 1) {
#############################################################################
## data       => contains scale | degree | stars | speaker.p | prior.p | cost
## calcFN     => without cost = 1, with cost = 2
## alpha      => set alpha level

  results <-
    data.frame(scale = c(),
               stars = c(),
               degree = c(),
               preds = c())

  scales <- unique(data$scale)
  for (s in scales) {
    force(
      mat <- subset(data, scale  == s) %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
#         mutate(hi1 = hi1 / sum(hi1),
#                hi2 = hi2 / sum(hi2),
#                mid = mid / sum(mid),
#                low1 = low1 / sum(low1),
#                low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, mid, low1, low2)
      )
    stars <- subset(data, scale == s)$stars
    degrees <- as.character(subset(data, scale == s)$degree)
    currCosts <- unique(subset(data, scale == s) %>% select(degree, cost))
    priors <- data %>% select(c(stars, prior.p)) %>% head(., n=5)

@@

Data
```{r}
# priors
# -------
unifPriors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
empPriors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
salience <- c(0.22, 0.08, 0.05, 1, 1)
saliencePriors <- data.frame(stars=1:5, prior.p = (salience / sum(salience)))
salienceCosts <- data.frame(degree = c("low2", "low1", "mid", "hi2", "hi1"),
                            cost = -1 * (salience / sum(salience)), stringsAsFactors = FALSE)
emptyCosts <- data.frame(degree = c("low2", "low1", "mid", "hi2", "hi1"),
                            cost = rep(0.0, 5), stringsAsFactors = FALSE)
yelpPriors <- data.frame(stars=1:5,
                         prior.p =
                           c(0.090190981, 0.105689431, 0.162783722, 0.304569543, 0.336766323))

# speaker
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
# listener
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")

costs <- salienceCosts
priors <- unifPriors

data <- left_join(speakerL12, costs) %>%
  left_join(., priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, prior.p, cost) %>%
  group_by(scale)
```

Explore
```{r}
test <- runFull2(data, calcFn = 2, alpha = 20)

predDf <- test
getOverallCor <- function(predDf, listenerDf) {
  combinedDf <- data.frame(scale = c(), degree = c(), speaker.p = c(), prior.p = c(), cost = c(), preds = c())
  for (s in unique(predDf$scale)) {
    if (s == "some_all") {
      tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "mid")) %>%
        mutate(degree = ifelse(degree == "hi1", "hi", "low"))  
    } else {
      tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
        mutate(degree = ifelse(degree == "hi1", "hi", "low"))  
    }
    combinedDf <- rbind(combinedDf, tempDf)
  }
  results <- merge(x = combinedDf,
                   y = listenerDf,
                   by = c("degree", "stars", "scale"),
                   all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))

  cor(results$preds, results$listener.p)
}
getOverallCor(test, listener)

tuneAlphas <- function(alphas = seq(1, 12), humanDf) {
  sapply(alphas, function(i) {
    predDf <- runFull2(data, calcFn = 2, alpha = i)
    getOverallCor(predDf, humanDf)
    })
}
tuneAlphas(humanDf = listener)
```

```{r}

normedVec(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
#normedVec(subset(test, scale == "good_excellent" & degree == "mid")$preds)
sum(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
dfPreds <- test
dfHuman <- listener
name <- "good_excellent"
getCorrs <- function(dfPreds, dfHuman, name) {
  predSubset <- subset(dfPreds, scale == name & (degree == "hi1" | degree == "hi2"))
  predSubset$degree = with(predSubset,
                           ifelse(degree == "hi1", "hi", "low"))
  humanSubset <- subset(dfHuman, scale == name)
  joinedDf <- merge(x = predSubset,
                    y = humanSubset,
                    by = c("degree", "stars", "scale"),
                    all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))
  cor(joinedDf$preds, joinedDf$listener.p)
}

>>>>

New Prior incorporation
```{r}
speaker.inf3 = function(d, alpha, cost = 0) {
  exp(-1 * alpha * (-log(d) - cost))
  # exp(alpha * (log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd3 = function(currRating, currDegree, m, alpha, costs, allDegrees) {
  currCost <- subset(costs, degree == currDegree)$cost
  numerator = speaker.inf3(m[[currDegree]][currRating], alpha, currCost)
  # normalize over other words
  normalize <- sum(sapply(seq(allDegrees), FUN = 
                            function(i) {
                              speaker.inf(m[[allDegrees[i]]][currRating], alpha,
                                          subset(costs, degree == allDegrees[i])$cost)}
                          ))
  return(numerator / normalize)
  # return(numerator)
}
# old !!!!
# speaker.lhd = function(rating, degree, m, alpha) {
#   numerator = speaker.inf(m[rating, degree], alpha)
#   normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
#   return(numerator / normalize)
# }

@@

Run RSA - five item scales
```{r}
computePosteriors <- c(norm.post, norm.post3)
# runFull() 
# ---------
# data       => contains scale | degree | stars | speaker.p | prior.p | cost
# calcFN     => without cost = 1, with cost = 2
# alpha      => set alpha level
runFull <- function(data, calcFn = 1, alpha = 1, numAlts = 5) {
  results <-
    data.frame(scale = c(),
               stars = c(),
               degree = c(),
               preds = c())

  scales <- unique(data$scale)
  for (s in scales) {
      if (numAlts == 5) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
#           mutate(hi1 = hi1 / sum(hi1),
#                  hi2 = hi2 / sum(hi2),
#                  mid = mid / sum(mid),
#                  low1 = low1 / sum(low1),
#                  low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, mid, low1, low2)
      } else if (numAlts == 4) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, low1, low2)
      } else if (numAlts == 2) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2)) %>%
          select(hi1, hi2)
      }
    stars <- subset(data, scale == s)$stars
    degrees <- as.character(subset(data, scale == s)$degree)
    currCosts <- unique(subset(data, scale == s) %>% select(degree, cost))
    priors <- data %>% select(c(stars, prior.p)) %>% head(., n=5)

@@

Data set-up
```{r}
# All scales studied
# ------------------
scales <- c("good_excellent",
            "liked_loved",
            "palatable_delicious",
            "memorable_unforgettable",
            "some_all")
# Priors
# ------
unifPriors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
empPriors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
yelpPriors <- data.frame(stars=1:5,
                         prior.p =
                           c(0.090190981, 0.105689431, 0.162783722, 0.304569543, 0.336766323))
# Cost / rewards
# --------------
emptyCosts <- data.frame(degree = c("low2", "low1", "mid", "hi2", "hi1"),
                            cost = rep(0.0, 5), stringsAsFactors = FALSE)
salienceCosts <- read.csv("~/Desktop/Projects/scalar_implicature/models/saliences.csv")
for (s in scales) {
  salienceCosts[salienceCosts$scale == s, ]$cost <-
    -1 * normedVec(salienceCosts[salienceCosts$scale == s, ]$cost)
}
# salience <- c(0.22, 0.08, 0.05, 1, 1)
# saliencePriors <- data.frame(stars=1:5, prior.p = (salience / sum(salience)))
# salienceCosts <- data.frame(degree = c("low2", "low1", "mid", "hi2", "hi1"),
#                             cost = -1 * (salience / sum(salience)), stringsAsFactors = FALSE)

# Speaker
# -------
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
# Listener
# --------
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")

costs <- salienceCosts
priors <- unifPriors
data <- left_join(speakerL12, costs) %>%
  left_join(., priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, prior.p, cost) %>%
  group_by(scale)

# run fns()
checkSaliences()
tuneAlphas(modelData = data, humanDf = listener)
```

Explore
```{r}
# matchModelHumandData()
# -----------------------
matchModelHumanData <- function(predDf, listenerDf, study) {
  combinedDf <- data.frame(scale = c(),
                           degree = c(),
                           speaker.p = c(),
                           prior.p = c(),
                           cost = c(),
                           preds = c())

  for (s in unique(predDf$scale)) {
    if (s == "some_all") {
      tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "mid")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
    } else {
      tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
    }
    combinedDf <- rbind(combinedDf, tempDf)
  }

  results <- merge(x = combinedDf,
                   y = listenerDf,
                   by = c("degree", "stars", "scale"),
                   all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))

  results
}
# mergedData <- matchModelHumanData(test, listener)

# Alpha Tuning - maximixize correlation
# -------------------------------------
tuneAlphas <- function(modelData, humanDf, alphas = seq(1, 12), numAlts = 5){
  sapply(alphas, function(i) {
    predDf <- runFull(modelData, calcFn = 2, alpha = i, numAlts = numAlts)
    mergedData <- matchModelHumanData(predDf, listener)
    cor(mergedData$preds, mergedData$listener.p)
  })
}
tuneAlphas(modelData = data, humanDf = listener)

checkSaliences <- function() {
  predDf <- runFull(data, calcFn = 2, alpha = 1, numAlts = 5)
  mergedData <- matchModelHumanData(predDf, listener)
  cor(mergedData$preds, mergedData$listener.p)
}
checkSaliences()
```

Visualize
```{r}
mergedData <- matchModelHumanData(runFull(data, calcFn = 2, alpha = 7, numAlts = 5), listener)
plotData <- mergedData
qplot(stars, listener.p, col=degree,
      data=plotData,
      main=paste("title")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = preds), lty = 4)

correlationScatterPlot <-
  ggplot(mergedData, aes(x = listener.p, y = preds, col=stars)) +
  geom_point() + 
  stat_smooth(method="lm")

```{r}
normedVec(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
#normedVec(subset(test, scale == "good_excellent" & degree == "mid")$preds)
sum(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
dfPreds <- test
dfHuman <- listener
name <- "good_excellent"
getCorrs <- function(dfPreds, dfHuman, name) {
  predSubset <- subset(dfPreds, scale == name & (degree == "hi1" | degree == "hi2"))
  predSubset$degree = with(predSubset,
                           ifelse(degree == "hi1", "hi", "low"))
  humanSubset <- subset(dfHuman, scale == name)
  joinedDf <- merge(x = predSubset,
                    y = humanSubset,
                    by = c("degree", "stars", "scale"),
                    all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))
  cor(joinedDf$preds, joinedDf$listener.p)
}
